---
title: "Hw03_couBC"
author: "CouBC"
date: '2018-10-01'
output: github_document
---
#Assignment 3:
Loading tidyverse and gapminder.

```{r}
library(gapminder)
library(tidyverse)
```

#Task 1:
Get the maximum and minimum of GDP per capita for all continents

```{r}
glimpse(gapminder)
```

At first I tried this - but this is not helpful at all as you would have to scroll through all the rows to try to find the max and min manually. 

```{r}
gapminder %>%
  arrange(gdpPercap)
  
```

I tried a couple of different combinations with max(gdpPercap) and min(gdpPercap) and pmax(x) and pmin(x) but I couldn't get it to work telling me there were errors in my argument. I looked at Jenny's tutorial on one-table verbs (http://stat545.com/block010_dplyr-end-single-table.html) which looked at max and min life expectancie:   summarize(min_lifeExp = min(lifeExp), max_lifeExp = max(lifeExp)) and tried that for gdpPercap. Voila:


```{r}
gapminder %>%
  group_by(continent) %>% 
  summarise(max_gdpPercap = max(gdpPercap), min_gdpPercap = min(gdpPercap))
```


```{r}
gapminder %>%
  group_by(continent) %>% 
  summarize(max(gdpPercap), min(gdpPercap))
```

Because it may be more meaningful to know max and min gdpPercap for a particular year, in the next section, I filtered to 2007:

```{r}
gapminder %>%
  group_by(continent) %>% 
  filter(year == 2007) %>% 
  summarise(max_gdpPercap = max(gdpPercap), min_gdpPercap = min(gdpPercap))
```

In order to visualize this in a graph:

```{r}
ggplot(gapminder, aes(x=continent, y=gdpPercap)) +geom_boxplot() + scale_y_log10() + ggtitle("Comparing GDP per capita across continents")
```

Only looking at year 2007 data for gdp Per capita across continents:

```{r}
gapminder_2007 <- gapminder %>%
  filter(year == 2007)

ggplot(gapminder_2007, aes(x = continent, y = gdpPercap)) +
  geom_boxplot() +
  scale_y_log10() + ggtitle("Comparing GDP per capita across continents")
```

            

#Task 2
Look at spread of GDP per capita within the continents.


```{r}
gapminder %>% 
  group_by(continent) %>% 
  filter(year == 2007) %>% 
  select(country, gdpPercap) %>% 
  summarize(sd_gdpPercap = sd(gdpPercap))
```



These are my different attempts at visualising the spread of gdp Per capita across the continents in the year 2007. It looks like Asia has the most spread of gdpPercap while Oceania has the least.


```{r}

gapminder_2007 <- gapminder %>%
  filter(year == 2007, continent != "Oceania")

ggplot(gapminder_2007, aes(x = year, y = gdpPercap, color = continent)) +
  geom_violin() + facet_wrap(~continent) + 
  geom_jitter(alpha =.2) +
  scale_y_log10() +
  ggtitle("Comparing spread of GDP per capita across continents")
```

I tried looking at spread in 2007 using boxplots which is useful for showing more specific information (IQR, range) about the the spread of gdpPercap between each continent.

```{r}
gapminder_2007 <- gapminder %>%
  filter(year == 2007)
ggplot(gapminder_2007, aes(x = continent, y = gdpPercap)) +
  geom_boxplot() +
   scale_y_log10() +
    ggtitle("Comparing GDP per capita across continents using boxplots")


```

            
#Task 3
How is life expectancy changing over time on different continents?

Need to figure out how to find the difference in lifeExp between the latest data point (2007) and first data point (1952). 

After trying several unsuccessful ways of creating a new variable that looks at change in lifeExp, I looked at the participation repo for cm_08.

```{r}
gapminder %>% 
    group_by(continent) %>% 
    transmute(lifeExp_change = lifeExp - lag(lifeExp)) %>% 
    summarize(median_change = median(lifeExp_change, na.rm = T)) %>%
  knitr::kable()

```

```{r}
gapminder %>% 
    group_by(continent, country) %>% 
    mutate(lifeExp_change = lifeExp - lag(lifeExp)) %>% 
    ggplot(aes(year, lifeExp_change)) +
    geom_point() + facet_wrap(~continent, scales = "free_y")
  
```


```{r}
ggplot(gapminder, aes(x = year, y = lifeExp, color = continent)) +
  geom_point() +
  facet_grid(~continent) +
  ggtitle("Change in Life Expectancy over 55 years by Continent") 
```

I found the above graph too busy and difficult to understand/visualize the continental data. I tried another way of looking at it. I adapted the code from: https://www.mrozinski.com.pl/posts/introduction-to-tidyverse/




```{r}
by_year_continent <- gapminder %>%
      group_by(continent, year) %>% 
      summarize(mdn_lifeExp = median(lifeExp))

ggplot(by_year_continent, aes(x=year, y=mdn_lifeExp, color=continent)) + 
      geom_point() + 
      expand_limits(y=0) +
 ggtitle("Change in Life Expectancy over 55 years by Continent")
```

