---
title: "Hw03_couBC"
author: "CouBC"
date: '2018-10-01'
output: 
  html_document: 
    keep_md: yes
---
#Assignment 3:
Loading tidyverse and gapminder.

```{r}
library(gapminder)
library(tidyverse)
library(dplyr)
```

#Attempt on the first task:
Get the maximum and minimum of GDP per capita for all continents

```{r}
View(gapminder)
glimpse(gapminder)
```

At first I tried this - but this is not helpful at all as you would have to scroll through all the rows to try to find the max and min by hand. 

```{r}
gapminder %>%
  group_by(continent) %>%
  arrange(gdpPercap)
  
```

I tried a couple of different combinations with max(gdpPercap) and min(gdpPercap) and pmax(x) and p(min) but I couldn't get it to work telling me there were errors in my argument. I looked at Jenny's tutorial on one-table verbs (http://stat545.com/block010_dplyr-end-single-table.html) which looked at max and min life expectancie:   summarize(min_lifeExp = min(lifeExp), max_lifeExp = max(lifeExp)) and tried that for gdpPercap. Voila:

```{r}
gapminder %>%
  group_by(continent) %>% 
  summarise(max_gdpPercap = max(gdpPercap), min_gdpPercap = min(gdpPercap))
```

Because it may be more meaningful to know max and min gdpPercap for a particular year, in the next section, I filtered to 2007:

```{r}
gapminder %>%
  group_by(continent) %>% 
  filter(year == 2007) %>% 
  summarise(max_gdpPercap = max(gdpPercap), min_gdpPercap = min(gdpPercap))
```

In order to visualize this in a graph:

```{r}
ggplot(gapminder, aes(x=continent, y=gdpPercap)) +geom_boxplot() + scale_y_log10() + ggtitle("Comparing GDP per capita across continents")
```

Only looking at year 2007 data for gdp Per capita across continents:

```{r}
gapminder_2007 <- gapminder %>%
  filter(year == 2007)

ggplot(gapminder_2007, aes(x = continent, y = gdpPercap)) +
  geom_boxplot() +
  scale_y_log10() + ggtitle("Comparing GDP per capita across continents")
```

            
            
            

#Task 2
Look at spread of GDP per capita within the continents.

            
#Task 3
How is life expectancy changing over time on different continents?


